### 2016-05-20 old_classification.tex
update table (autogenerated)

\section{Summary of models}

Throu visual inspection of the most popular models generated for each candidate and data generated by those models we classified the candidates in four categories.
The data available is shown in \figref{SW58}.
\todo{better explain criterias?}
The following sections list all the models for all the categories, each with a short explanation of the reasons.
Additionally, selected models are presented in figures for demonstration.

\subsection{plausible}

The plausible candidates fail in one category, but are quite convincing in all the others.
One particular reason for a candidate to end up in this category is that is has extended arcs, what is hard to modell with the current version of the program.
Thats why we recomend them for a following up study, for example for taking spectras or other follow up obersatories and for other modelling programs for a try.

\lenstitle{SW01 (ASW0004dv8)}
  This is some text that says something more than this line.

\lenstitle{SW08 (ASW00099ed)}
  Simple lens with two images;
  model has minor arthefacts.

\lenstitle{SW11 (ASW0002qtn)}
  Simple small arc;
  model nedes minor refinement;
  quite low mass ratio

\lenstitle{SW12 (ASW0003wsu)} 
  faint arc, bright overall situation, no counter image;
  has good model;
  low Mstel and low Mlens

\lenstitle{SW21 (ASW0004m3x)} 
  clear arc, no counter image; inbetween two external sources;
  model needs work, predicts additional images. play with external mass / sheer?
  Mass ratio low.

\lenstitle{SW27 (ASW0006jh5)} 
  (really plausible?)
  unclear situation, but arc is visible;
  model shows extended arc not there in org image;
  mass ratio ok

\lenstitle{SW31 (ASW00021r0)} 
  (really plausible?)
  noisy situation, unclear images, potential external mass interfering;
  model ok'ish, hard to say;
  high mass ratio

\lenstitle{SW33 (ASW0003s0m)}
  clear picture, two image situation;
  model simple, but ok. arc too much extended;
  masses in the middle, but rather high ratio

\lenstitle{SW36 (ASW000096t)} 
  noisy picture, many possible external masses, faint images;
  model predicts additional images, needs more work, arc too extended;
  masses look ok

\lenstitle{SW38 (ASW0009cp0)} 
  quite clean picture, two bright minima, counter image hidden;
  model fine
  really hi stellar mass

\lenstitle{SW43 (ASW0001c3j)} 
  clean situation, faint arc'ish images;
  model needs work, shows additional images, strange arc (LHD);
  very high Mstel

\lenstitle{SW45 (ASW00024id)}
  ring like structure, very clean
  usual problems with rings, otherwise model quite ok actually;
  very high Mstel

\lenstitle{SW46 (ASW00024q6)} 
  yellow'ish arc, two image configuration;
  model ok;
  masses ok

\lenstitle{SW52 (ASW0006a07)} 
  (no imgs)

\lenstitle{SW54 (ASW0007sez)} 
  white lens, fuzzy ring like structure;
  model really good, given ring like;
  masses ok

\lenstitle{SW58 (ASW0007iwp)} 
  nice arc, no counter image;
  model prediction really good;
  masses ok
  good example of plausible

\subsubsection{grouplens}

\lenstitle{SW36 (ASW000096t)}
(like 4dv8 collective modelling)

\subsection{unclear}

This categy contains the ones that are hard to say.
They might look like lenses visually, but the modelling failed or prediced additional images.

\lenstitle{SW04 (ASW0009cjs)} 
  noisy field, possible group lens, but images not typical? (no arclike structure);
  model fail (input errors);
  Mlens too high (10e13)

\lenstitle{SW06 (ASW0008swn)}
  green images, elongated lens, images on strange positions?;
  model ok'ish, but mass distr. doesn't reflect visuals;
  masses ok

\lenstitle{SW07 (ASW0007e08)} 
  (no img)

\lenstitle{SW13 (ASW00047ae)} 
  croweded situation unclear setup;
  model failed, predicts additional images;
  masses ok

\lenstitle{SW15 (ASW0004nan)} 
  no counter image (should be visible)?
  model ok'ish;
  Mlens on the lower end

\lenstitle{SW17 (ASW0005rnb)} 
  two image setup, but arcish structure of images missing;
  model failed, predicts additional images, rednering failed;

\lenstitle{SW19 (ASW0001ld7)} 
  asymmetry / really extended arc and no counter image rather no lensing configuration;
  rendering ok'ish, model predicts additionla images;
  Mstel and Mlens way too low (e10 / e11)

\lenstitle{SW24 (ASW00050sk)} 
  unclear / noisy situation, shape of arc strange;
  model ok;
  mass ration close to unity

\lenstitle{SW32 (ASW0004iye)} 
  (no img)

\lenstitle{SW35 (ASW0004wgd)} 
  hard to say, doesn't look convincing, but neither totally off
  modell ok;
  masses as well;

\lenstitle{SW44 (ASW0002k40)} 
  (no img);
  very old model;
  hi mass ratio

\lenstitle{SW47 (ASW0003r6c)} 
  two images, no arcish structre.. too far away from lens?;
  model predicts additional images;
  quite high mass ratio

\lenstitle{SW51 (ASW0006e0o)} 
  (no img)

\lenstitle{SW56 (ASW0007pga)} 
  very ellyptical lens, images not really arcish;
  model failed, predicts additional images;
  masses ok

\lenstitle{SW57 (ASW0008pag)} 
  point mass / group lens interfering? symertry of arc off;
  model failed completly;
  Mlens too high

\subsection{convincing}

The canidates that looked the most promising and left little doubt about being lenses were grouped into the category ``convincing''.
In Figures \ref{fig:SW02} to \ref{fig:SW29} we show all the convincing candidates.
Following we list the parameters, obtained applying the technique described in section~\ref{sec:parameter}

\lenstitle{SW02 (ASW000619d)}
Nice arc with counter image;
additional mass at 12 o'clock doesn't seem to influence the image.
\params{1.04}{0.32}{-5.18}

\lenstitle{SW05 (ASW0007k4r)}
The model predicts very steep central mass profile, and thus a very high lensing mass;
looks otherwise very convincing.
\params{3.80}{0.15}{52.50}

\lenstitle{SW09 (ASW0002asp)}
\params{1.12}{0.00}{88.39}

\lenstitle{SW28 (ASW0007xrs)}
Faint, but clearly visible arc and counter image;
but model sfits nicely
\params{1.04}{0.33}{66.63}

\lenstitle{SW29 (ASW0008qsm)}
Faint close arc, but reconstruion works very well;
parameterisation of ellypticity fails.
\params{0.73}{0.00}{249.80}

\subsection{doubtful}

\lenstitle{SW10 (ASW0002bmc)} 
  very elliptical setup, unlikley to be a lens;
  model failed totally

\lenstitle{SW16 (ASW0009bp2)} 
  very faint images

\lenstitle{SW18 (ASW0007hu2)} 
  model failed completly

\lenstitle{SW20 (ASW0002dx7)} 
  faint white images, rather substuctre of ``lens'' ??

\lenstitle{SW22 (ASW0009ab8)} 
  very faint images, group configuration;
  model predicts additional images;
  mass ration close to unity

\lenstitle{SW23 (ASW0003r61)} 
  (no img)

\lenstitle{SW26 (ASW0005ma2)} 
  similar to SW10

\lenstitle{SW34 (ASW00051ld)} 
  (no img)

\lenstitle{SW41 (ASW0008xbu)} 
  similar to SW10. very ellyptical lens

\lenstitle{SW42 (ASW00096rm)} 
  ring like structure, but very strange distribution;
  model not too bad, brightnesses off;
  stellar mass way too lower

\lenstitle{SW53 (ASW00070vl)} 
  images don't look liked lensed;
  model predicts additional images

\begin{figure}
  \inclfig{SW42}{ASW00096rm_4Q3YCEWGLN}
  \caption{outlier because very odd mass ratio}
  \label{fig:SW42}
\end{figure}

\subsection{missing}

Finally this category lists all the candidates, that miss redshift measurements of the lens and thus further analysis is difficult.

\lenstitle{SW03 (ASW0006mea)}

\lenstitle{SW14 (ASW0004xjk)}

\lenstitle{SW25 (ASW00007mq)}

\lenstitle{SW30 (ASW0002p8y)}

\lenstitle{SW37 (ASW00086xq)}

\lenstitle{SW39 (ASW0005qiz)}

\lenstitle{SW40 (ASW0008wmr)}

\lenstitle{SW48 (ASW0000g95)}

\lenstitle{SW49 (ASW00007ls)}

\lenstitle{SW50 (ASW00008a0)}

\lenstitle{SW55 (ASW0007t5y)}

\lenstitle{SW59 (ASW00085cp)}