### 2016-12-12 parts/spl.tex
edits

\section{Characterising the lens models}\label{sec:diag}