### 2017-02-15 parts/figures.tex
edits

\begin{figure*}
\inclGrid{spl-input}
\caption{Nine of the lens candidates marked up with spaghetti
  diagrams.  Red, blue and green dots are proposed locations for
  maxima, minima and saddle points of the arrival time.  The curves
  help guide the placement of the dots, but their precise appearance
  has no significance.  This selection includes the best-modelled
  systems, but also one case (SW57 at upper right) of unsuccessful
  modelling.  Since the modelling process is collaborative among the
  volunteers, with anyone welcome to contribute new models or modify
  existing ones, there are variant spaghetti diagrams for all the
  modelled systems.  The online supplement displays all the models
  presented for discussion during this work.
\label{fig:splinput}}
\end{figure*}

\begin{figure*}
\inclGrid{arrival_spaghetti}
\caption{Arrival-time surfaces for models of the systems from
  Figure~\ref{fig:splinput}.  These are cleaned-up versions of the
  original spaghetti diagrams.  The registration differs slightly from
  Figure~\ref{fig:splinput}, but the coloured dots represent exactly
  the sky positions specified in the earlier figure.  The orange
  contours only qualitatively resemble the earlier pink curves, as
  they are now precise saddle-point contours from lens models.
  \label{fig:arriv}}
\end{figure*}

\begin{figure*}
\inclGrid{nsynth}
\caption{Synthetic images of the systems from
  Figures~\ref{fig:splinput}, derived from the lens models.  The
  reconstructed lensed features are keep the Space~Warps false-colour
  scheme from Figures~\ref{fig:splinput}.  The rest has been changed
  to black-and-white.
  \label{fig:synth}}
\end{figure*}

\begin{figure*}
\inclGrid{kappa_map_interpol}
\caption{Mass distribution $\kappa$ for which the results
  Figures~\ref{fig:arriv}--\ref{fig:synth} were derived.
\label{fig:kappa}}
\end{figure*}

\begin{figure*}
\inclGrid{kappa_encl}
\caption{Cumulative circular-averages of the mass maps from
  \ref{fig:kappa}, with uncertainties.  More precisely, we show the
  enclosed mass within a given projected radius, expressed as the mean
  $\kappa$ with a given number of arcsec from the centre of the
  lensing galaxy.  The orange bands refer to the full ensemble of mass
  maps for the models, while the red curves show the ensemble
  averages.  The dashed vertical line indicates the notional Einstein
  radius, or where the mean enclosed $\kappa$ is unity.  The short
  vertical arrows marks the positions of the images (maxima, saddle
  points and minima).
  \label{fig:encl}}
\end{figure*}

\begin{figure*}
\includegraphics[width=\linewidth]{img/mlens_vs_mstel_one/mstel_vs_mtot_one}
\caption{Total mass in the model against the estimated stellar mass,
  alongside the values for the whole sample.  (The labelled orange
  bars are the systems shown in detail in
  Figures~\ref{fig:splinput}--\ref{fig:encl}.)  The lower-right
  shaded region is unphysical according to the stellar-population
  models, because it gives $M<\Mstel$. The upper-left shaded region is
  unphysical according to abundance matching, because it gives
  $M>\Mhalo$.  That is to say, the unshaded region is
  $0<\haloindex<1$. \label{fig:stelmass}}
\end{figure*}

\begin{table*}
  \caption{Categorisation of SW models}
  \label{tab:models}
  \include{table}
\end{table*}